Integer-N Frequency Synthesizer Designed for Settling Time Compliance

### 2025-06-03 CaseStudies/Example3_3_IntegerNSynthesizerDesignedforSettlingTime/article/article.qmd
asdkjfasldkjfhasf

---
title: Integer-N Frequency Synthesizer Designed for Settling Time Compliance
subtitle: Case Study - Example 3.3 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design of an integer-N frequency synthesizer, focusing on the theoretical and practical aspects of system dimensioning. The objective is to ensure precise frequency generation within a specified range, while meeting constraints related to frequency resolution and settling time. The design methodology involves the definition of key dynamic parameters, such as natural frequency, damping factor, and loop bandwidth, based on continuous-time control theory. A voltage-controlled oscillator (VCO) is configured to provide the required tuning range, accounting for process variations and voltage supply limitations. Additionally, a second-order passive loop filter is designed to ensure stability and appropriate dynamic response. The overall approach demonstrates a systematic framework for the design of simple frequency synthesizers.
keywords: 
    - Integer-N PLL 
    - Settling Time
    - Continuous-Time Analysis
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

from IPython.display import display, Markdown

def format_graph(value):
    abs_value = abs(value)
    if abs_value >= 1e9:
        result = f"{value / 1e9:.2f} G"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:.2f} M"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:.2f} k"
    elif abs_value >= 1:
        result = f"{value * 1e0:.2f} "
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:.2f} m"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:.2f} μ"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:.2f} n"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:.2f} p"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:.2f} f"
    else:
        result = f"{value:.0e} "

    return result
```

@@

from IPython.display import display, Markdown

def format(value, decimals=2):
    abs_value = abs(value)
    fmt = f".{decimals}f"

    if abs_value >= 1e9:
        result = f"{value / 1e9:{fmt}}\\,\\text{{G}}"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:{fmt}}\\,\\text{{M}}"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:{fmt}}\\,\\text{{k}}"
    elif abs_value >= 1:
        result = f"{value * 1e0:{fmt}}\\,"
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:{fmt}}\\,\\text{{m}}"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:{fmt}}\\,\\mu"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:{fmt}}\\,\\text{{n}}"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:{fmt}}\\,\\text{{p}}"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:{fmt}}\\,\\text{{f}}"
    else:
        result = f"{value:.0e}\\,"

    return Markdown(f"${result}$")

```{python}
#| echo: false
import math
from IPython.display import display, Markdown

def format_with_pi_latex(value):
    pi_multiple = value / math.pi
    abs_pi_multiple = abs(pi_multiple)

    if abs_pi_multiple >= 1e9:
        result = f"{pi_multiple / 1e9:.2f}\\pi\\,\\text{{G}}"
    elif abs_pi_multiple >= 1e6:
        result = f"{pi_multiple / 1e6:.2f}\\pi\\,\\text{{M}}"
    elif abs_pi_multiple >= 1e3:
        result = f"{pi_multiple / 1e3:.2f}\\pi\\,\\text{{k}}"
    elif abs_pi_multiple >= 1:
        result = f"{pi_multiple:.2f}\\pi"
    elif abs_pi_multiple >= 1e-3:
        result = f"{pi_multiple * 1e3:.2f}\\pi\\,\\text{{m}}"
    elif abs_pi_multiple >= 1e-6:
        result = f"{pi_multiple * 1e6:.2f}\\pi\\,\\mu"
    elif abs_pi_multiple >= 1e-9:
        result = f"{pi_multiple * 1e9:.2f}\\pi\\,\\text{{n}}"
    elif abs_pi_multiple >= 1e-12:
        result = f"{pi_multiple * 1e12:.2f}\\pi\\,\\text{{p}}"
    elif abs_pi_multiple >= 1e-15:
        result = f"{pi_multiple * 1e15:.2f}\\pi\\,\\text{{f}}"
    else:
        result = f"{pi_multiple:.0e}\\pi"

    return Markdown(f"${result}$")

# Exemplo de uso:

@@

f_o = f0=2.4e9
f=2.4835e9
t_s = 225e-6
f_r = f_ref = 1e6
```

This report presents the design of an integer-N frequency synthesizer operating from `{python} format(f0)`Hz to `{python} format(f, 4)`Hz. The synthesizer is required to achieve frequency settling within a time of $t_s=$ `{python} format(t_s)`s. Given that the channel spacing is `{python} format(f_ref)`Hz and an integer-N architecture is employed, the channel spacing is equal to the reference frequency. This example is based on the methodology and parameters provided in @rogers2006IntegratedCircuitDesign.

# Design Methodology

## Frequency Settling and Loop Dynamics

To meet the settling time requirement, the natural frequency $\omega_n$ of the loop must be determined. Using an equation discussed in a previous report, the relationship between settling time, damping factor $\zeta$, and natural frequency is given by:

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

@@

zeta=0.707
omega_n__t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
omega_n=omega_n__t_s/t_s
f_n=omega_n/(2*math.pi)
```

Assuming an arbitrary value of $\zeta=$ `{python} zeta`, the resulting natural frequency is $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s, which corresponds to a frequency $f_n=$ `{python} format(f_n)`Hz.

From the natural frequency and damping factor, the 3-dB bandwidth $\omega_{3\text{dB}}$ can be calculated as:

$$
\omega_{3\text{dB}}=\omega_n\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}
$$

This results in a bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz. Given that $f_\text{ref}\gg f_{3\text{dB}}$, the discretization effects due to sampling are negligible, and a continuous-time analysis remains valid for this system.

```{python}
#| echo: false
N0=f0/f_ref
N=f/f_ref
```

Since the architecture is integer-N and the reference frequency is $f_\text{ref}=$ `{python} format(f_ref)`Hz, the division ratio $N$ must range from `{python} round(N0, 0)` to `{python} round(N, 0)`.

## Voltage-Controlled Oscillator (VCO)

f_range=f-f0
tuning_range_percent = 9.6 
tuning_range = f0/tuning_range_percent
Vdd=2
Kvco = 250e6 # Hz/V
Kvco_ang = Kvco*2*math.pi
```

The VCO must span a tuning range of `{python} format(f_range)`Hz. To account for process variations, an additional tuning margin is considered. A total tuning range of approximately `{python} round(tuning_range_percent, 0)`% of the center frequency, or about `{python} format(tuning_range)`Hz, is recommended.

Assuming a supply voltage of `{python} format(Vdd)`V, a VCO gain of `{python} format(Kvco)`Hz/V (or `{python} format(Kvco_ang)`rad/s/V) is expected. The free-running frequency of the VCO is set at the minimum frequency. The center frequency $f_c$ is given by:

$$
f_c=\frac{f+f_0}{2}
$$

@@

f_c=(f+f0)/2
fvco=f_min=f_c-tuning_range
f_max=f_c+tuning_range

f_example_1=2.401e9
```

This results in $f_c=$ `{python} format(f_c)`Hz. Therefore, the minimum and maximum VCO frequencies are $f_\text{min}=$ `{python} format(f_min)`Hz and $f_\text{max}=$ `{python} format(f_max)`Hz, respectively. The free-running frequency is set as $f_\text{VCO}=f_\text{min}$.

For instance, to synthesize a frequency of `{python} format(f_example_1, 3)`Hz, the required control voltage $V_c$ is computed using:

$$
\begin{split}
f_o&=f_\text{VCO}+K_\text{VCO}\cdot V_c\implies\\
V_c&=\frac{f_o-f_\text{VCO}}{K_\text{VCO}}
\end{split}
$$

@@

V_c_example_1=(f_example_1-fvco)/Kvco
```

This yields a control voltage of $V_c=$ `{python} format(V_c_example_1)`V.

## Loop filter

As previously noted, continuous-time analysis is applicable. Standard second-order loop filter equations are used. The first capacitor $C_1$ can be calculated from:

$$
C_1=\frac{I_\text{CP}K_\text{VCO}}{2\pi\cdot N\omega_n^2}
$$

@@

C_1=5e-9
I_CP_ang=C_1*2*math.pi*N0*omega_n**2/Kvco_ang
I_CP=C_1*2*math.pi*N0*omega_n**2/Kvco

Since the charge pump current $I_\text{CP}$ is unknown, a value of $C_1=$ `{python} format(C_1)`F is assumed, leading to a calculated $I_\text{CP}=$ `{python} format(I_CP_ang)`A. The resistor $R_1$ is then given by:

$$
R_1=2\zeta\sqrt{\frac{2\pi\cdot N}{IK_\text{VCO}C_1}}
$$

R_1=2*zeta*math.sqrt((2*math.pi*N0)/(I_CP*Kvco*C_1))

Thus, the resistor value is $R=$ `{python} format(R_1)`$\Omega$.

Using a common design heuristic, the second capacitor $C_2$ is:

$$
C_2=\frac{C_1}{10}
$$

C_2=C_1/10 
```

Which results in $C_2=$ `{python} format(C_2)`F.

# Parameters

## Maximum Phase Error Instant

Given the value of $\omega_n$, the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

$$
t_m=\frac{1}{\omega_n\sqrt{1-\zeta^2}}\tan^{-1}\left(\frac{\sqrt{1-\zeta^2}}{\zeta}\right)
$$ {#eq-instant-maximum-frequency-step}

@@

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

[Eq. @eq-instant-maximum-frequency-step] represents the instant at which the phase error reaches its maximum value.

## Maximum Normalized Phase Error

The maximum normalized phase error, $\theta_{e_\text{m,n}}$, is given by:

$$
\begin{split}
\theta_{e_\text{m,n}}&=\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t_m\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt_m}
\end{split}
$$

@@

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 2 * math.pi
```

And applying $t_m$ to this equation, the maximum normalized phase error in this case is approximately `{python} round(theta_e_m_n, 2)`.

## Maximum Allowed Frequency Step

This normalized value can be used in the following non-normalized equation for the phase error:

$$
\begin{split}
\theta_e(t)&=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
\end{split}
$$

The goal is to find the maximum input frequency step, $\Delta\omega_m$, such that the maximum phase error does not exceed $\theta_{e_m}=2\pi$. Using the normalized result:

$$
\begin{split}
\theta_{e_m}&=\frac{\Delta\omega_m}{\omega_n}\theta_{e_\text{m,n}}\implies\\
\Delta\omega_m&=\frac{\theta_{e_m}\omega_n}{\theta_{e_\text{m,n}}}
\end{split}
$$

@@

Substituting the calculated values yields:

- $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s
- $\Delta f_m=$ `{python} format(step_m_lin)`Hz

Hence, the maximum input frequency step that still allows theoretical predictability is `{python} format(step_m_lin)`Hz. If the input frequency step exceeds this limit, the PLL may lose lock and exhibit cycle slipping (more details are provided in the next section).

## Settling Time Estimation

The settling time, $t_s$, required for the system to reach a steady state after a frequency step can be approximated by:

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta*omega_n)*(-1)
```

This results in a settling time of $t_s=$ `{python} format(t_s)`s.

# Simulation

<!--```{python}
#| echo: false

# Parâmetros iniciais
f_r = 10e6
f_o = 50e6
N = 5
KVCO = 10e6
I = 1e-3
Kphase = I / (2 * np.pi)
A0 = 1
K = A0 * Kphase * KVCO / N
```-->

# Parâmetros iniciais
Kphase = I_CP / (2 * np.pi)
A0 = 1
K = A0 * Kphase * Kvco / N0
```

<!--```{python}
#| echo: false

from scipy import signal

omega_n = omega_bw / np.sqrt(1 + 2 * zeta**2 + np.sqrt(4 * zeta**4 + 4 * zeta**2 + 2))
f_n = omega_n/2/np.pi

# Parâmetros do PLL
C1 = K / omega_n**2
C2 = C1 / 10
R1 = 2 * zeta * omega_n / K
```-->

@@

from scipy import signal
import numpy as np

# Numerador e denominador do compensador F(s)
# F(s) = (s*C1*R1 + 1) / (s^2*C1*C2*R1 + s*(C1 + C2))
F_num = [C_1 * R_1, 1]
F_den = [C_1 * C_2 * R_1, C_1 + C_2, 0]

F_sys = signal.TransferFunction(F_num, F_den)

# F(s) * K / s = G(s)
G_num = np.polymul(F_num, [K])
G_den = np.polymul(F_den, [1, 0])  # Multiplica por s

# H(s) = G(s) / (1 + G(s))
H_num = G_num
H_den = np.polyadd(G_den, G_num)

# Criando o sistema de tempo contínuo
H_sys = signal.TransferFunction(H_num, H_den)

@@

```{python}
#| echo: false
#| label: fig-pll-step-response
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

@@

# Tempo de simulação
t = np.linspace(0, 300e-6, 100000)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = step_m_ang_meas
u = np.ones_like(t) * step_amplitude

@@

# Gerar gráfico
plt.figure(figsize=(8, 5))
plt.plot(t_out * 1e6, y_out, label='Step Response')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Step')
plt.axhline(final_value + tol, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(final_value - tol, color='red', linestyle='--', linewidth=0.8)
plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Graphical settling time: {ts_graphical*1e6:.5f} us')
plt.xlabel('Time [us]')
plt.ylabel('Phase [rad]')
plt.title('Step Response with Settling Time (99%)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-vc-step-response
#| fig-cap: "Voltage control step response of the designed PLL."
#| fig-align: center

@@

# Numerador e denominador da função de transferência (V_C / ω_R)
num = [(N0 * omega_n_meas**2 / Kvco) * (2 * zeta_meas / omega_n_meas), N0 * omega_n_meas**2 / Kvco]
den = [1, 2 * zeta_meas * omega_n_meas, omega_n_meas**2]

# Criando o sistema (sem multiplicar por s — assume entrada ω_R = Δω/s → degrau em ω_R)
system = TransferFunction(num, den)

# Tempo de simulação
t = np.linspace(0, 300e-6, 1000)

# Resposta ao degrau
t, vc_response = step(system, T=t)
vc_response *= step_m_lin

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin) * N0 - f_o) / Kvco
tol = 0.01 * final_value  # 1% de tolerância
upper_bound = final_value + tol
lower_bound = final_value - tol

@@

>>>>

---
title: Integer-N Frequency Synthesizer Designed for Settling Time Compliance
subtitle: Case Study - Example 3.3 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    Example 3.3.
keywords: 
    - Integer-N PLL 
    - Settling Time
    - Continuous-Time Analysis
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

@@

from format_utils import format, format_graph, format_no_markdown, format_with_pi_latex

@@

f_o = fo_1 = 2.4e9
omega_o_1 = 2 * np.pi * fo_1

fo_2 = 2.4835e9
omega_o_2 = np.pi * 2 * fo_2

f_r = f_ref = 1e6
omega_r = np.pi * 2 * f_r

t_s = 225e-6

zeta=0.707
```

omega_n__t_s = math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
omega_n = omega_n__t_s/t_s
f_n = omega_n/(2*math.pi)

@@

N1 = fo_1/f_ref
N2 = fo_2/f_ref
```

@@

f_range = fo_2 - fo_1
tuning_range_percent = 9.6 
tuning_range = fo_1 / tuning_range_percent
Vdd=2
KVCO = Kvco = 250e6 # Hz/V
KVCO_ang = Kvco_ang = Kvco*2*math.pi
```

@@

f_c=(fo_1 + fo_2) / 2
fvco = f_min = f_c - tuning_range
f_max=f_c+tuning_range
```

@@

import math 
import numpy as np

C_1 = C1 = 5e-9
A0 = 1

I_CP_ang = C_1 * 2 * math.pi * N1 * omega_n**2 / Kvco_ang / A0
I = I_CP = C_1 * 2 * math.pi * N1 * omega_n**2 / Kvco / A0
K_theta = Kphase = I_CP / (2 * np.pi)

K = A0 * Kphase * Kvco / N1
```

@@

R_1 = R1 = 2 * zeta * omega_n / K
C_2 = C2 = C_1 / 10 
```

@@

t_m = 1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

@@

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 1.6255 * math.pi
```

@@

@@

def poly_to_str(coeffs):
    terms = []
    degree = len(coeffs) - 1
    for i, coef in enumerate(coeffs):
        power = degree - i
        if abs(coef) < 1e-12:  # ignora coeficientes muito pequenos
            continue
        term = ""
        coef_str = format_eng(coef)
        if not (abs(coef - 1) < 1e-12 and power != 0):
            term += coef_str
        if power > 0:
            term += "s"
            if power > 1:
                term += f"^{power}"
        terms.append(term)
    return " + ".join(terms)
```

# Numerador e denominador do compensador F(s)
# F(s) = (s*C1*R1 + 1) / (s^2*C1*C2*R1 + s*(C1 + C2))
F_num = [C1 * R1, 1]
F_den = [C1 * C2 * R1, C1 + C2, 0]

F_sys = signal.TransferFunction(F_num, F_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = F_sys.num[0][0] if isinstance(F_sys.num[0], (list, tuple)) else F_sys.num
den = F_sys.den[0][0] if isinstance(F_sys.den[0], (list, tuple)) else F_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-filter-numeric}}"))

@@

```{python}
#| echo: false
# F(s) * K / s = G(s)
G_num = np.polymul(F_num, [K])
G_den = np.polymul(F_den, [1, 0])  # Multiplica por s

G_sys = signal.TransferFunction(G_num, G_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = G_sys.num[0][0] if isinstance(G_sys.num[0], (list, tuple)) else G_sys.num
den = G_sys.den[0][0] if isinstance(G_sys.den[0], (list, tuple)) else G_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$G(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-open-loop-numeric}}"))
```

```{python}
#| echo: false
# H(s) = G(s) / (1 + G(s))
H_num = G_num
H_den = np.polyadd(G_den, G_num)

# Criando o sistema de tempo contínuo
H_sys = signal.TransferFunction(H_num, H_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = H_sys.num[0][0] if isinstance(H_sys.num[0], (list, tuple)) else H_sys.num
den = H_sys.den[0][0] if isinstance(H_sys.den[0], (list, tuple)) else H_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$H(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-closed-loop-numeric}}"))
```

# Agora temos o denominador completo
# O_sys = H_sys * N
# Para isso, O_sys = (H_num * N) / (H_den)
num_O = np.convolve(H_sys.num, [N1])
den_O = np.convolve(H_sys.den, [1])

# Criar a função de transferência final
O_sys = signal.TransferFunction(num_O, den_O)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = O_sys.num[0][0] if isinstance(O_sys.num[0], (list, tuple)) else O_sys.num
den = O_sys.den[0][0] if isinstance(O_sys.den[0], (list, tuple)) else O_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$O(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-output-phase-numeric}}"))
```

# Kvco/s => [Kvco] / [1, 0]
Kvco_over_s = ([Kvco], [1, 0])

# F_sys => ([1], [1, 1]) já está
# Multiplicando numeradores e denominadores via convolução
num_den = np.convolve([1/N1], Kvco_over_s[0])
num_den = np.convolve(num_den, F_sys.num)
num_den = np.convolve(num_den, [K_theta])
num_den = np.convolve(num_den, [A0])

den_den = np.convolve(Kvco_over_s[1], F_sys.den)

# Agora temos o denominador completo
# E_sys = H_sys / (den)
# Para isso, E_sys = (H_num * den_den) / (H_den * den_num)
num_E = np.convolve(H_sys.num, den_den)
den_E = np.convolve(H_sys.den, num_den)

# Criar a função de transferência final
E_sys = signal.TransferFunction(num_E, den_E)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = E_sys.num[0][0] if isinstance(E_sys.num[0], (list, tuple)) else E_sys.num
den = E_sys.den[0][0] if isinstance(E_sys.den[0], (list, tuple)) else E_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$E(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-phase-error-numeric}}"))
```

# Vc_sys = H_sys / ((1/N)*(Kvco/s))
#Kvco_over_s = ([Kvco], [1, 0])
Kvco_over_s = ([Kvco], [1])
den_num = np.convolve([1/N1], Kvco_over_s[0])
den_den = Kvco_over_s[1]

num_Vc = np.convolve(H_sys.num, den_den)
den_Vc = np.convolve(H_sys.den, den_num)

Vc_sys = signal.TransferFunction(num_Vc, den_Vc)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = Vc_sys.num[0][0] if isinstance(Vc_sys.num[0], (list, tuple)) else Vc_sys.num
den = Vc_sys.den[0][0] if isinstance(Vc_sys.den[0], (list, tuple)) else Vc_sys.den

# Exibe como equação no Jupyter
display(Markdown(f"$$V_C(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-vctrl-vco-numeric}}"))
```

sim_stop_time = 500e-6
sim_points = 10000
step_amplitude = f_r
```

```{python}
#| echo: false
#| label: fig-step-freq-h
#| fig-cap: "Frequency step response of the closed-loop transfer function."
#| fig-align: center

@@

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
u = np.ones_like(t) * step_amplitude

@@

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback frequency')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference frequency')
plt.axhline(final_value + tol, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(final_value - tol, color='red', linestyle='--', linewidth=0.8)
plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Settling time: {ts_graphical*1e6:.5f} µs')
plt.xlabel('Time (µs)')
plt.ylabel('Frequency (Hz)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-ramp-phase-h
#| fig-cap: "Phase response to a ramp input in the closed-loop transfer function."
#| fig-align: center

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = omega_r * t
u = np.ones_like(t) * step_amplitude

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference Phase')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-ramp-phase-h-o
#| fig-cap: "Comparison of phase responses to a ramp input in two different functions."
#| fig-align: center

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = omega_r * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)
t_out, y_out2, _ = signal.lsim(O_sys, U=u, T=t)

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, y_out2, label='Output Phase')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference Phase')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-ramp-phase-step-freq-h
#| fig-cap: "Phase response to a ramp with an abrupt frequency change (frequency step)."
#| fig-align: center

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Frequências
f_base = omega_r  # 10 MHz
f_step = step_m_ang_meas  # step de 10 MHz
t_step = sim_stop_time / 2  # instante do step

# Entrada em fase: integral da frequência
phi_in = np.where(
    t < t_step,
    f_base * t,
    f_base * t + f_step * (t - t_step)
)

# Simulação da resposta
t_out, y_out, _ = signal.lsim(H_sys, U=phi_in, T=t)

### --- Gráfico ---
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, phi_in, linestyle=':', color='black', label='Reference Phase')

plt.axvline(t_step * 1e6, color='purple', linestyle='--', linewidth=1.2, label='Step Frequency Instant')

plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-ramp-phase-h-phase-error
#| fig-cap: "Response to a ramp input with phase error."
#| fig-align: center

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Entrada: rampa em fase (step de 10 MHz em frequência)
# Fase = ∫ frequência ⇒ rampa
freq_step = omega_r  # Step de 10 MHz em frequência
phi_in = freq_step * t  # Rampa de fase

# Simulação da resposta do sistema à rampa
t_out, phi_out, _ = signal.lsim(H_sys, U=phi_in, T=t)

# Erro de seguimento
tracking_error = phi_in - phi_out

# Plotagem
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, phi_in, linestyle=':', color='black', label='Reference Phase')
plt.plot(t_out * 1e6, phi_out, label='Feedback Phase')
plt.plot(t_out * 1e6, tracking_error, label='Phase Error')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-step-freq-e
#| fig-cap: "Phase error response to a frequency step input."
#| fig-align: center

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = step_m_ang_meas * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(E_sys, U=u, T=t)

# Cálculo do valor máximo (pico)
max_value = np.max(np.abs(y_out))
tol = 0.01 * max_value  # 1% do valor de pico (centrado em zero)

# Novo cálculo do tempo de acomodação (settling time)
within_bounds = np.abs(y_out) <= tol
ts_graphical_peak = 0
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical_peak = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Phase Error')

# Linha no valor máximo (positivo)
plt.axhline(max_value, color='blue', linestyle=':', linewidth=1.2, label=f'Max Value')

# Linhas horizontais em ±1% do valor de pico (em torno de zero)
plt.axhline(tol, color='red', linestyle='--', linewidth=0.8, label=f'±1% Tolerance')
plt.axhline(-tol, color='red', linestyle='--', linewidth=0.8)

# Marcação do novo tempo de acomodação
plt.axvline(ts_graphical_peak * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Settling Time: {ts_graphical_peak*1e6:.3f} μs')

plt.xlabel('Time (µs)')
plt.ylabel('Phase Error (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-step-freq-e-norm
#| fig-cap: "Normalized phase error response to a frequency step input."
#| fig-align: center

# Tempo de simulação
t = np.linspace(0, 8/omega_n_meas, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(E_sys, U=u, T=t)

y_out *= omega_n_meas

# Cálculo do valor máximo (pico)
max_value = np.max(np.abs(y_out))
tol = 0

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * omega_n_meas, y_out, label='Step Response')

# Linha no valor máximo (positivo)
plt.axhline(max_value, color='blue', linestyle=':', linewidth=1.2, label=f'Max Value: {max_value:.3f}')

# Linhas horizontais em ±1% do valor de pico (em torno de zero)
plt.axhline(tol, color='black', linestyle=':')

plt.xlabel('Time (µs)')
plt.ylabel('Phase Error norm. (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```{python}
#| echo: false
#| label: fig-step-freq-vc
#| fig-cap: "Control voltage response to a frequency step input."
#| fig-align: center

@@

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Resposta ao degrau
t, vc_response = step(Vc_sys, T=t)
vc_response *= step_m_lin_meas

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin) * N1 - fo_1) / KVCO
#final_value = 0.5
tol = 0.01 * final_value  # 1% de tolerância
upper_bound = final_value + tol
lower_bound = final_value - tol

@@

# Janela de amostras para verificar se permanece dentro da tolerância
window_size = 3*150  # ~5% da simulação
